---
title: "Peggy Adams Shelter Flow Viz Analysis"
author: "VCAP"
date: "2024-10-6"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, warning = FALSE, echo = FALSE}
#libraries
library(tidyverse)
library(ggalluvial)
library(kableExtra)
library(readr)

setwd("C:/Users/john/Documents/GitHub_Data/VCAP-Shelter-Flow-Data/PeggyAdmas")

peggy_adams_final_csv <- read_csv("final_peggy_adams_dat.csv")
peggy_adams_final_rds <- readRDS("final_peggy_adams_dat.rds")

```


```{r, echo = FALSE, warning = FALSE}

#final data prep

# turn data to quintiles or quartile
numToQuintile <- function(x,n=5){
  if(n==5){
    return (
      case_when(
        x>=0 & x<=0.2 ~ 1,
        x>0.2 & x<=0.4 ~ 2,
        x>0.4 & x<=0.6 ~ 3,
        x>0.6 & x<=0.8 ~ 4,
        x>0.8 & x<=1 ~ 5
      ))
  }
  if(n==4){
    return (
      case_when(
        x>=0 & x<=0.25 ~ 1,
        x>0.25 & x<=0.5 ~ 2,
        x>0.5 & x<=0.75 ~ 3,
        x>0.75 & x<=1 ~ 4
      ))
  }
}

# svi names to theme
svi_theme_names <- c(
  'RPL_THEME1' = 'Socioeconomic Status',
  'RPL_THEME2' = 'Household Characteristics',
  'RPL_THEME3' = 'Racial & Ethnic Minority Status',
  'RPL_THEME4' = 'Housing Type & Transportation',
  'RPL_THEMES' = 'SVI Overall Ranking'
)

animal_adoptions_SVI <- peggy_adams_final_rds %>% 
  filter(`Outcome Type` == "Adoption")
non_adoption_SVI <- peggy_adams_final_rds %>% 
  filter(`Outcome Type` != "Adoption")

intake_census_tracts <- animal_adoptions_SVI %>%
  
  select(intake_census_tract, RPL_THEME1_INT, RPL_THEME2_INT, RPL_THEME3_INT, RPL_THEME4_INT, RPL_THEMES_INT, E_HH_INT) %>% 
  rename("GEOID" = "intake_census_tract",
         "RPL_THEME1" = "RPL_THEME1_INT",
         "RPL_THEME2" = "RPL_THEME2_INT",
         "RPL_THEME3" = "RPL_THEME3_INT",
         "RPL_THEME4" = "RPL_THEME4_INT",
         "RPL_THEMES" = "RPL_THEMES_INT",
         "E_HH" = "E_HH_INT")

outcome_census_tracts <- animal_adoptions_SVI %>% 
  select(outcome_census_tract, RPL_THEME1_OUT, RPL_THEME2_OUT, RPL_THEME3_OUT, RPL_THEME4_OUT, RPL_THEMES_OUT, E_HH_OUT) %>% 
  rename("GEOID" = "outcome_census_tract",
         "RPL_THEME1" = "RPL_THEME1_OUT",
         "RPL_THEME2" = "RPL_THEME2_OUT",
         "RPL_THEME3" = "RPL_THEME3_OUT",
         "RPL_THEME4" = "RPL_THEME4_OUT",
         "RPL_THEMES" = "RPL_THEMES_OUT",
         "E_HH" = "E_HH_OUT")

SVI_breakdown <-  rbind(intake_census_tracts, outcome_census_tracts) %>% 
  distinct(GEOID, .keep_all = TRUE)
  

```

## Title 1

**Figure 1.** SVI Quintile distribution at the Census Tract level
```{r echo=FALSE, fig.height=4, fig.width=9, warning=FALSE, dpi=400}

SVI_breakdown %>%
  mutate(across(starts_with('RPL'), function(x) numToQuintile(x))) %>%  # SVIs to Qrtiles
  select(GEOID, starts_with('RPL')) %>%
  pivot_longer(-GEOID, names_to = 'svi_group', values_to = 'quintile') %>%
  mutate(svi_group = svi_theme_names[svi_group],
         quintile = as.character(quintile)) %>%
  drop_na() %>% 
  ggplot(aes(x=svi_group, fill=quintile)) +
  geom_bar(position = "fill") +
  labs(y='Percent', x='SVI Theme', fill="Quintile", 
       title=sprintf('Breakdown of %s SVI %s Rankings', "Peggy Adams","Quintile"))+
  coord_flip()+
  scale_fill_viridis_d(option = "plasma", direction = 1, alpha = 0.6) +
  theme(axis.text.y = element_text(size=12),
        legend.position = "top")+
  #theme(legend.position = "none")+
  scale_y_continuous(labels = scales::percent) +
  geom_text(
    aes(label=paste0(round(100*signif(after_stat(count) / tapply(after_stat(count), after_stat(x), sum)[as.character(after_stat(x))], digits=3),1),'%')),
    stat="count",
    position=position_fill(vjust=0.5))


```


## Title 2

**Figure 2.** Flow of animals from intake to adoption by SVI quintile.
```{r echo=FALSE, fig.height=7, fig.width=10, warning=FALSE, dpi=400}


animal_adoptions_SVI %>% 
  mutate(across(starts_with('RPL'), function(x) numToQuintile(x))) %>%  # SVIs to Qrtiles
  count(RPL_THEMES_INT, RPL_THEMES_OUT) %>%
  mutate(RPL_THEMES_INT=factor(RPL_THEMES_INT)) %>% #, labels = c("1", "2", "3", "4","5"))) %>%
  mutate(RPL_THEMES_OUT=factor(RPL_THEMES_OUT)) %>% # , labels = c("1", "2", "3", "4","5"))) %>%
  # viz
  ggplot(aes(axis1 = RPL_THEMES_INT, axis2 = RPL_THEMES_OUT, y = n)) +
  geom_alluvium(aes(fill = RPL_THEMES_INT) , curve_type = "quintic") +
  geom_stratum(alpha=0.3) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), nudge_x=-0.04) +
  scale_x_discrete(expand = c(.1, .1), limits = c(sprintf("Intake SVI %s","Quintile"), sprintf("Outcome SVI %s","Quintile")))+
  theme(axis.text.x = element_text(size=14))+
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_y_continuous(labels = NULL, breaks = NULL) + labs(y = NULL)+
  geom_text(stat = "stratum", aes(label = paste0('(',scales::percent(after_stat(prop),accuracy = .1),')')),
            nudge_x = 0.03)+
  theme(legend.position = "none")
  

```

## Title 3

91.5% of "Transfer In" intakes came from SVI quintile 5. Of the 1,881 animals transferred in, 1,689 came from 1 location. 

**Figure 3.** Flow of animals from intake to adoption by SVI quintile, divided by intake type
```{r, echo = FALSE, warning = FALSE, fig.width = 12, fig.height=7, fig.align='center', dpi=400}

animal_adoptions_SVI %>%
  #mutate(src_intake_type=ifelse(src_intake_type=='Return','Owner/Guardian Surrender',src_intake_type)) %>%
  #filter(src_intake_type %in% c('Owner/Guardian Surrender','Stray','Seized / Custody')) %>%
  #  ***intake plus species combo for sue
  #mutate(src_intake_type=paste(species,src_intake_type)) %>%
  #  ***dog size filters
  #filter(species=='Dog') %>% mutate(size_group=ifelse(size_group=='Extra Large','Large',size_group)) %>% 
  #filter(size_group!="") %>%
  #  age 
  # mutate(age_group=case_when(
  #   age_group=='Unknown' ~ 'Adult',
  #   age_group=='Senior' ~ 'Adult',
  #   age_group=='Neonate' ~ 'Youth',
  #   age_group=='Juvenile' ~ 'Youth',
  #   TRUE ~ age_group
  # )) %>% 
  #filter(age_group=='Adult') %>% # for the figure of size without Youths
  #mutate(age_group = paste(species, age_group)) %>% #count(age_group)
  mutate(across(starts_with('RPL'), function(x) numToQuintile(x))) %>%  # SVIs to Qrtiles
  # ADD N TO FACET AND CHOOSE FACET VARIABLE (twice)
  group_by(`Intake Type`) %>% mutate(n_group=n()) %>% mutate(`Intake Type`=paste0(`Intake Type`,' (n = ',formatC(n_group, big.mark = ','),')')) %>% ungroup() %>%
  count(`Intake Type`, RPL_THEMES_INT, RPL_THEMES_OUT) %>%
  mutate(RPL_THEMES_INT=factor(RPL_THEMES_INT)) %>% #, labels = c("1", "2", "3", "4","5"))) %>%
  mutate(RPL_THEMES_OUT=factor(RPL_THEMES_OUT)) %>% #, labels = c("1", "2", "3", "4","5"))) %>%
  # viz
  ggplot(aes(axis1 = RPL_THEMES_INT, axis2 = RPL_THEMES_OUT, y = n)) +
  geom_alluvium(aes(fill = RPL_THEMES_INT) , curve_type = "quintic") +
  geom_stratum(alpha=0.3) +
  # CHOOSE FACET HERE TOO
  facet_wrap(~`Intake Type`, scales='free_y')+   # FACET
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),nudge_x=-0.05) +
  geom_text(stat = "stratum", aes(label = paste0('(',scales::percent(after_stat(prop),accuracy = .1),')')),
            nudge_x = 0.04)+
  scale_x_discrete(expand = c(.1, .1), limits = c(sprintf("Intake SVI %s","Quintile"), sprintf("Outcome SVI %s","Quintile")))+
  theme(axis.text.x = element_text(size=14))+
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_y_continuous(labels = NULL, breaks = NULL) + labs(y = NULL)+
  theme(strip.text = element_text(size=12))+
  theme(legend.position = "none")


```


## Title 4

**Figure 4.** Flow of animals from intake to adoption by SVI quintile, rotating SVI themes

```{r, warning = FALSE, echo = FALSE, fig.width = 10, fig.height=8, dpi=400}

animal_adoptions_SVI %>%
  mutate(id = row_number()) %>% 
  mutate(across(starts_with('RPL'), function(x) numToQuintile(x))) %>%  # SVIs to Qrtiles
  select(id, matches('RPL_THEME[1-4].*')) %>%
  pivot_longer(-id, names_to=c('which_svi','INT_OUT'),names_pattern = '(.*)_(.*)',
               values_to = 'quintile') %>%
  pivot_wider(id_cols=c(id, which_svi),
              names_from = 'INT_OUT', values_from = 'quintile') %>%
  count(which_svi, RPL_THEMES_INT=`INT`,RPL_THEMES_OUT=OUT) %>% # confirm it works filter(which_svi=='RPL_THEME1') %>% pull(n) %>% sum()
  mutate(which_svi = svi_theme_names[which_svi]) %>%
  mutate(RPL_THEMES_INT=factor(RPL_THEMES_INT)) %>% #, labels = c("1", "2", "3", "4"))) %>%
  mutate(RPL_THEMES_OUT=factor(RPL_THEMES_OUT)) %>% # , labels = c("1", "2", "3", "4"))) %>%
  # viz
  ggplot(aes(axis1 = RPL_THEMES_INT, axis2 = RPL_THEMES_OUT, y = n)) +
  geom_alluvium(aes(fill = RPL_THEMES_INT) , curve_type = "quintic") +
  geom_stratum(alpha=0.3) +
  facet_wrap(~which_svi, scales='free_y')+   # FACET
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),nudge_x=-0.05) +
  geom_text(stat = "stratum", aes(label = paste0('(',scales::percent(after_stat(prop),accuracy = .1),')')),
            nudge_x = 0.04)+
  scale_x_discrete(expand = c(.1, .1), limits = c(sprintf("Intake SVI %s","Quintile"), sprintf("Outcome SVI %s","Quintile")))+
  theme(axis.text.x = element_text(size=11))+
  theme(strip.text = element_text(size=14))+
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_y_continuous(labels = NULL, breaks = NULL) + labs(y = NULL)+
  theme(legend.position = "none")

```

## Title 5

**Figure 5.** Flow of animals from intake to adoption by SVI quintile, divided by age and species

```{r, echo = FALSE, warning = FALSE, fig.width = 10, fig.height = 8, dpi=400}

animal_adoptions_SVI %>%
  #  age 
  mutate(age_group = paste(Species, Age_Group_at_Outcome)) %>% #count(age_group)
  mutate(across(starts_with('RPL'), function(x) numToQuintile(x))) %>%  # SVIs to Qrtiles
  # ADD N TO FACET AND CHOOSE FACET VARIABLE (paste twice)
  group_by(age_group) %>% mutate(n_group=n()) %>% mutate(age_group=paste0(age_group,' (n = ',formatC(n_group, big.mark = ','),')')) %>% ungroup() %>%
  count(age_group, RPL_THEMES_INT, RPL_THEMES_OUT) %>%
  mutate(RPL_THEMES_INT=factor(RPL_THEMES_INT)) %>% #, labels = c("1", "2", "3", "4","5"))) %>%
  mutate(RPL_THEMES_OUT=factor(RPL_THEMES_OUT)) %>% #, labels = c("1", "2", "3", "4","5"))) %>%
  # viz
  ggplot(aes(axis1 = RPL_THEMES_INT, axis2 = RPL_THEMES_OUT, y = n)) +
  geom_alluvium(aes(fill = RPL_THEMES_INT) , curve_type = "quintic") +
  geom_stratum(alpha=0.3) +
  # CHOOSE FACET HERE TOO
  facet_wrap(~age_group, scales='free_y')+   # FACET
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),nudge_x=-0.05) +
  geom_text(stat = "stratum", aes(label = paste0('(',scales::percent(after_stat(prop),accuracy = .1),')')),
            nudge_x = 0.04)+
  scale_x_discrete(expand = c(.1, .1), limits = c(sprintf("Intake SVI %s","Quintile"), sprintf("Outcome SVI %s","Quintile")))+
  theme(axis.text.x = element_text(size=11))+
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_y_continuous(labels = NULL, breaks = NULL) + labs(y = NULL)+
  theme(strip.text = element_text(size=12))+
  theme(legend.position = "none")


```

## Title 6

**Figure 6.** Intakes and adoptions per 100 households within SVI quintiles.

```{r, warning=FALSE, echo = FALSE, dpi=400}

min_per_tract <- 10

SVI_breakdown %>%
  # include only tracts with some intakes
 filter(GEOID %in% (c(animal_adoptions_SVI$intake_census_tract,animal_adoptions_SVI$outcome_census_tract,non_adoption_SVI$intake_census_tract) %>%
     table %>% data.frame() %>% filter(Freq>=min_per_tract) %>% pull(1) %>% as.character())) %>% #
 mutate(across(starts_with('RPL'), function(x) factor(numToQuintile(x)))) %>%
  select(GEOID, RPL_THEMES, hh=E_HH) %>%
 left_join(animal_adoptions_SVI %>% select(GEOID=intake_census_tract) %>% 
             rbind(non_adoption_SVI %>% select(GEOID=intake_census_tract)) %>%
             count(GEOID), by='GEOID') %>% # join in 
 left_join(animal_adoptions_SVI %>% count(GEOID=outcome_census_tract), by='GEOID', suffix=c('_in','_out')) %>%
 replace(is.na(.),0) %>% # if no intake or outcome insert 0
 group_by(RPL_THEMES) %>% 
 summarize(total_hh=sum(hh), sum_in=sum(n_in), sum_out=sum(n_out)) %>%
 mutate(in_per_k = (sum_in)/total_hh*1000, out_per_k=sum_out/total_hh*1000) %>%
  # viz - barplot
  select(RPL_THEMES, Intakes=in_per_k, Adoptions=out_per_k) %>%
  pivot_longer(-RPL_THEMES, names_to = 'variable',values_to = 'value') %>%
  ggplot(aes(x=reorder(RPL_THEMES,desc(RPL_THEMES)), y=value, fill=variable)) +
  geom_bar(stat='identity', position = "dodge") +
  scale_fill_manual(values = c("#6A60D1", "#D8576BFF"))+
  coord_flip()+
  labs(x='SVI Quintile', y='Animals per 1000 Households', fill='') +
  geom_text(
    aes(label=round(value)),
    size=4, position=position_dodge(.9),
    hjust=1.25)+
  theme(text = element_text(size=12),
        legend.position = "top")


```

## Title 7

**Figure 7.** Intakes and adoptions per 1000 households within SVI quintiles, divided by intake type.

```{r, echo = FALSE, warning = FALSE, dpi=400}
# same figure but now faceted by intake type
SVI_breakdown %>%
  filter(GEOID %in% (c(animal_adoptions_SVI$intake_census_tract,animal_adoptions_SVI$outcome_census_tract,non_adoption_SVI$intake_census_tract) %>%
      table %>% data.frame() %>% filter(Freq>=min_per_tract) %>% pull(1) %>% as.character())) %>%
  mutate(across(starts_with('RPL'), function(x) factor(numToQuintile(x)))) %>%
  select(GEOID, RPL_THEMES, hh=E_HH) %>%
  group_by(RPL_THEMES) %>% mutate(total_hh = sum(hh)) %>% ungroup() %>%
  left_join(animal_adoptions_SVI %>%  
              select(GEOID=intake_census_tract, `Intake Type`,Species) %>%
              rbind(non_adoption_SVI %>% select(GEOID=intake_census_tract, `Intake Type`,Species)) %>%
              count(GEOID, `Intake Type`) %>% 
              pivot_wider(names_from='Intake Type',values_from='n'),
              by='GEOID') %>% 
  left_join(animal_adoptions_SVI %>% 
              count(GEOID=outcome_census_tract, `Intake Type`) %>%
              pivot_wider(names_from='Intake Type',values_from='n'),
            by=c('GEOID'), suffix=c('_in','_out')) %>%
  replace(is.na(.),0) %>% # if no intake or outcome insert 0
  pivot_longer(-c(GEOID,RPL_THEMES,hh,total_hh), names_to=c('Intake Type','in_out'), 
              names_pattern = "(.*)_(.*)") %>%
  pivot_wider(names_from = 'in_out', values_from = 'value') %>%
  group_by(RPL_THEMES, total_hh, `Intake Type`) %>% 
  summarize(sum_in=sum(`in`), sum_out=sum(out)) %>%
  mutate(in_per_k = (sum_in)/total_hh*1000, out_per_k = (sum_out)/total_hh*1000) %>%
  ungroup() %>%
  select(`Intake Type`, RPL_THEMES, Intakes=in_per_k, Adoptions=out_per_k) %>%
  pivot_longer(-c(`Intake Type`,RPL_THEMES), names_to = 'variable',values_to = 'value') %>%
  ggplot(aes(x=reorder(RPL_THEMES,desc(RPL_THEMES)), y=value, fill=variable)) +
  geom_bar(stat='identity', position='dodge') +
  scale_fill_manual(values = c("#6A60D1", "#D8576BFF"))+
  facet_wrap(~`Intake Type`,nrow = 3)+
  coord_flip()+
  labs(x='SVI Quintile', y='Animals per 1000 Households', fill='') +
  theme(text = element_text(size=12),
        legend.position = "top")



```


## Title 8

There were 7 outcomes types listed, Adoption, Euthanasia, Feral/Wildlife, Return to Owner/Guardian, Transfer Out, Died, and Stolen/Lost. The last two were removed because they made up less than 1% of all outcomes.

**Figure 8.** Distribution of outcomes within intake SVI quintiles

```{r echo=FALSE, fig.height=4, fig.width=9, warning=FALSE, dpi=400}

peggy_adams_final_rds %>%
  filter(!`Outcome Type` %in% c("Died", "Stolen / Lost")) %>% 
  mutate(across(starts_with('RPL'), function(x) factor(numToQuintile(x)))) %>%  # SVIs to Qrtiles
  filter(!is.na(RPL_THEMES_INT)) %>%
  select(RPL_THEMES_INT, `Outcome Type`) %>%
  ggplot(aes(x=reorder(RPL_THEMES_INT,desc(RPL_THEMES_INT)), fill=`Outcome Type`)) +
  geom_bar(position = "fill") +
  labs(y='Percent', x='Outcome Type', fill=sprintf('Intake SVI\n%s',"Quintile"))+
  labs(y='Percent', x='In SVI Quintile ', fill='Outcome Type')+
  coord_flip()+
  scale_fill_viridis_d(option = "plasma", direction = 1, alpha = 0.6) +
  theme(axis.text.y = element_text(size=12))+
  theme(legend.text = element_text(size=12),
        legend.position = "top")+
  #theme(legend.position = "none")+
  scale_y_continuous(labels = scales::percent) +
  geom_text(
    aes(label=paste0(round(100*signif(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)], digits=3),1),'%')),
    stat="count", size=4,
    position=position_fill(vjust=0.5))

```

